---
title: "A tool box for a climate neutral housing sector: Climate Goals"
date: "`r Sys.Date()`"
output:
  pdf_document:
    number_sections: yes
  html_document:
    df_print: paged
    number_sections: yes
editor_options: 
  markdown: 
    wrap: 72
author:   
  - Anna Hornykewycz^[Institute for Comprehensive Analysis of the Economy (ICAE),
                  Johannes Kepler University, Linz, Austria]
  - Jakob Kapeller^[Institute for Socio-Economics, University of Duisburg-Essen, 
                  Germany and
                  Institute for Comprehensive Analysis of the Economy (ICAE),
                  Johannes Kepler University, Linz, Austria]
  - Jan David Weber^[Institute for Socio-Economics, University of Duisburg-Essen, 
                  Germany]
  - Bernhard Schütz^[Vienna Institute for International Economic Comparisions (wiiw)
                  and Institute for Socio-Economics, University of Duisburg-Essen, 
                  Germany]
  - Lukas Cserjan^[Institute for Comprehensive Analysis of the Economy (ICAE),
                  Johannes Kepler University, Linz, Austria]
---


# Defining relative reduction goals

Reduction goals are typically defined relative to a certain base year and set 
relative reduction targets for specific periods (for the \textit{German case} 
the base year is 1990, whereas targets are defined for milestones in 2030, 2040 
and 2050). Specifically, we formulate our climate goal estiamtes on the basis of 
the [Federal Climate Change Act (KSG)](https://www.gesetze-im-internet.de/englisch_ksg/index.html).
\footnote{We use Appendix I to construct the goal for 2030, while basing the 
goals for 2040 and 2050 on Appendix II.}

In doing so, we also adjust for current emissions to 
rescale targets to more accurately reflect contemporary needs for emission 
reduction. In this version, we use recent data to account for current emissions
(see [here](https://www.umweltbundesamt.de/presse/pressemitteilungen/klimaemissionen-sinken-2023-um-101-prozent) for more details); however, extrapolating the trend from 1990 onwards would 
give very similar results.

```{r climate_goals_parameters}
emissions_buildings_1990 = 210
goal_2030 = -(emissions_buildings_1990-67)/emissions_buildings_1990
goal_2040 = -0.88
goal_2050 = -1.00
emissions_buildings_2023 = 102 

target_2030 = emissions_buildings_1990 * (1 + goal_2030) 
target_2040 = emissions_buildings_1990 * (1 + goal_2040) 
target_2050 = emissions_buildings_1990 * (1 + goal_2050)

global_targetsinpercent <- c(0,
                      1 - target_2030 / emissions_buildings_2023,
                      1 - target_2040 / emissions_buildings_2023,
                      1 - target_2050 / emissions_buildings_2023)

```

# Extrapolation between targets

In this section we linearly extrapolate targets to generate an exhaustive 
time-series.

```{r calculating_time_series}
start_year = 2023
end_year = 2050

target_1 = 2030
target_2 = 2040
target_3 = end_year

globalscenario <- matrix(NA, end_year-start_year+1, 2) 
for (i in start_year:end_year) {
  if (i <= target_1) {
    globalscenario[i-start_year+1, ] <- c(i, (global_targetsinpercent[2] - 
                                    global_targetsinpercent[1]) / 
                                    (target_1 - start_year) * (i - start_year))
  } else if (i <= target_2) {
    globalscenario[i-start_year+1, ] <- c(i, global_targetsinpercent[2] + 
                                    (global_targetsinpercent[3] - 
                                     global_targetsinpercent[2]) / 
                                     (target_2 - target_1) * (i - target_1))
  } else if (i <= target_3) {
    globalscenario[i-start_year+1, ] <- c(i, global_targetsinpercent[3] + 
                                    (global_targetsinpercent[4] - 
                                     global_targetsinpercent[3]) / 
                                     (target_3 - target_2) * (i - target_2))
  }
}

globalscenario <- as.data.frame(globalscenario)
colnames(globalscenario) <- c("year", "rel_saving")
globalscenario$rel_reduction <- 1 - globalscenario$rel_saving
```

# Exporting the generated time series for further processing.

For file management, we make use of the `here`-package. PLease make sure to 
install this package and verify, whether the correct path is obtained.

```{r export_climate_goals, message=FALSE, results="hide"}
library(here)
here::here()
write.csv(globalscenario, 
          here::here("Intermediate_Results/global_buildings_climate_goals2040.csv"), 
          row.names = FALSE)

```

