---
title: 'A tool box for a climate neutral housing sector: Data Cleaning'
author:
- Anna Hornykewycz^[Institute for Comprehensive Analysis of the Economy (ICAE), Johannes
  Kepler University, Linz, Austria]
- "Jakob Kapeller^[Institute for Socio-Economics, University of Duisburg-Essen, Germany
  and Institute for Comprehensive Analysis of the Economy (ICAE), Johannes Kepler
  University, Linz, Austria]"
- "Jan David Weber^[Institute for Socio-Economics, University of Duisburg-Essen, Germany]"
- "Bernhard Schütz^[Vienna Institute for International Economic Comparisions (wiiw)
  and Institute for Socio-Economics, University of Duisburg-Essen, Germany]"
- Lukas Cserjan^[Institute for Comprehensive Analysis of the Economy (ICAE), Johannes
  Kepler University, Linz, Austria]
date: "`r Sys.Date()`"
output:
  pdf_document:
    number_sections: yes
  html_document:
    df_print: paged
    number_sections: yes
  word_document: default
---

# Preparations

## Load packages and set seed

This chunk gathers all relevant packages and sets one seed for all the random 
processes that follow.

```{r setup, warning=FALSE, message=FALSE}
library(here) 
library(tidyverse)
library(data.table)
library(ggplot2)
library(dplyr)
library(viridis)

set.seed(1301)
```
##  Import Climate Goals

Here, we import the climate goals.

```{r climate_coals}
climate_goals <- read.csv(here::here(
                 "Intermediate_Results/global_buildings_climate_goals2040.csv"))
```
## Setting basic parameters

```{r basic_parameters}
# basic policy parameters
rate_renovation <- 3
ec_goal <- 60

# basic technical parameters
apartment_per_building <- 6.85
```

## Importing a dataset

In this section micro-data on available residential units is imported. In 
addition, we identify all variables of interest, create a new variable for the 
type of building (flat vs. house), trim the dataset in several basic dimensions 
and translate key variables into english. 

When applying the code make sure that the files containing the raw data are in 
a folder "data" within the repository of this code file.

For \textit{Germany}, there is ample data available, so we restrict the units 
observed to those observed in housing ads on [immoscout24.de](https://www.immoscout24.de) 
published in 2020 or later.\footnote{In terms of concrete data vintages this 
implies using data files 7 to 14 for all relevant data groups (housing vs. 
flats, lease vs. sell).}

```{r dataimport, message=FALSE, warning=FALSE, results=FALSE}
v1 <- 7
v2 <- 14

path_str <- c(paste0("HKSUF",v1:v2), paste0("HMSUF",v1:v2), 
              paste0("WKSUF",v1:v2), paste0("WMSUF",v1:v2))

vars_of_interest <- c("ev_kennwert", "bef1", "spell", "letzte_modernisierung", 
                      "baujahr", "wohnflaeche", "immobilientyp", "zimmeranzahl",
                      "kategorie_Haus", "kategorie_Wohnung", 
                      "energieeffizienzklasse", "energieausweistyp", "anbieter",
                      "ejahr")

for (p in path_str){
  path <- here::here(paste0("Data/", p, ".csv"))
  raw <- data.table::fread(path, 
                          select=vars_of_interest,
                          colClasses=c(baujahr="character",
                                       wohnflaeche="character"))
  mod <- raw %>%  
    filter(spell == 1) %>% # this eliminates double entries
    filter(ejahr>2019) %>% 
    filter(kategorie_Haus!="Block of flats") %>% # eliminates unusable building type
    mutate(type = ifelse(grepl("H", p) == TRUE, "House", "Apartment"))
    # assign(paste0(p), mod) 
    # the line above imports individual files as separate data frames
  if(p == path_str[1]){
    immo_raw <- mod
  } else {
    immo_raw <- dplyr::bind_rows(immo_raw, mod)
  }
  print(p)
}

immo_raw <- immo_raw %>%
    mutate(living_area = as.double(wohnflaeche),
           ec_index = as.double(ev_kennwert),
           construction_year = as.double(baujahr),
           last_modernization = as.double(letzte_modernisierung)) %>%
    rename(
    efficiency_class_raw = energieeffizienzklasse,
    energy_certificate_type = energieausweistyp,
    provider = anbieter #,
    #living_area = wohnflaeche
  )
```

# Data processing

## Data cleaning

In a first step we eliminate entries where crucial data is missing and exclude 
implausible values for `living_area`.

```{r datacleaning}
na_strings <- c("Other missing", "Implausible value", "NO_INFORMATION", 
                "Not specified", NA, "NA")

immo_raw <- immo_raw %>%
  filter(!ec_index %in% na_strings) %>%
  filter(!bef1 %in% na_strings) %>%
  filter(!living_area %in% na_strings)

immo_raw <- immo_raw %>%
  filter(living_area > 20) %>%
  filter(living_area < 500)
```

## Creation of new variables

We create a new variable for the `efficiency_class` based on `ec_index` (which 
focuses more on final energy use than the information of efficiency classes as 
contained in our dataset for \textit{Germany}), as well as a variable for final 
energy use (`energy_use`). In doing we also transform some variables to numeric 
information.

Furthermore, we create a variable called `renovation weight`. This variable 
allows for taking the differences between houses and apartments into account 
and requires the average number of apartments per multi-unit building as an 
input. Finally, we collect different descriptions for available heating systems 
to arrive at a more consistently worded variable.

```{r newvariables, message=FALSE, warning=FALSE, results=FALSE}
immo_raw <- immo_raw %>%
  mutate(
    #living_area = as.double(living_area),
    efficiency_class = case_when(
      ec_index < 30 ~ "+",
      ec_index < 50 ~ "A",
      ec_index < 75 ~ "B",
      ec_index < 100 ~ "C",
      ec_index < 130 ~ "D",
      ec_index < 160 ~ "E",
      ec_index < 200 ~ "F",
      ec_index < 250 ~ "G",
      ec_index >= 250 ~ "H"
    ),
    energy_use = living_area * ec_index
  )

immo_raw <- immo_raw %>%
  mutate(
    renovation_weight = if_else(type == "Apartment", 
                                1 / apartment_per_building, 1))

other <- c("COMBINED_HEAT_AND_POWER_FOSSIL_FUELS" )
unclear <- c("HEAT_SUPPLY", "LOCAL_HEATING")
biomass <- c("WOOD","PELLET_HEATING","WOOD_CHIPS","BIO_ENERGY",
             "COMBINED_HEAT_AND_POWER_BIO_ENERGY",
             "COMBINED_HEAT_AND_POWER_REGENERATIVE_ENERGY")
heatingpump <- c("GEOTHERMAL", "ENVIRONMENTAL_THERMAL_ENERGY", 
                 "SOLAR_HEATING", "HYDRO_ENERGY", "WIND_ENERGY",
                 "COMBINED_HEAT_AND_POWER_RENEWABLE_ENERGY")
gas <- c("GAS", "LIQUID GAS", "NATURAL_GAS_LIGHT", "NATURAL_GAS_HEAVY")
district_heating <- c("DISTRICT_HEATING", "STEAM_DISTRICT_HEATING")
coal <- c("COAL", "COAL_COKE")
oil <- c("OIL")
electricity <-c("ELECTRICITY")

immo_raw <- immo_raw %>%
  mutate(
    heating_system = case_when(bef1 %in% electricity ~ "electricity",
                             bef1 %in% oil ~ "oil",
                             bef1 %in% coal ~ "coal",
                             bef1 %in% district_heating ~ "DH",
                             bef1 %in% gas ~ "gas",
                             bef1 %in% heatingpump ~ "HP & other renewables",
                             bef1 %in% biomass ~ "biomass",
                             bef1 %in% unclear ~ "unclear",
                             bef1 %in% other ~ "other"))

```

## Preliminary visualization of the dataset focusing on efficiency classes / energy use

Create a plot that shows the distribution of the different efficiency classes 
based on  `ec_index`.

```{r distributionplot, message=FALSE, warning=FALSE}
plot_all <- ggplot(immo_raw, aes(x = ec_index)) +
  geom_density(aes(colour = efficiency_class), alpha = .2, fill = "#FF6666",
               show.legend = TRUE) +
  labs(title = "All", x = "ec_index", y ="Density (accommodation units)") +
  xlim(0, 500) +
  ylim(0, 0.1) +
  theme_bw()

plot_all
```

Export the plot as a pdf file.

```{r distributionplotPDF, message=FALSE, warning=FALSE}
ggsave("../Graphs/plot_all.pdf", plot_all, width = 6, height = 4)

```

# Creating a representative sample

We acknowledge that our sample is not representative across all efficiency 
classes. Especially low efficiency classes are underrepresented in the available 
data. As the [real distribution](https://www.dena.de/fileadmin/dena/Publikationen/PDFs/2022/dena_Gebaeudereport_2023.pdf) 
of the efficiency classes, as well as the house vs flat distribution is known, 
we are able to create a dataset that is representative for these two key 
dimensions.\footnote{By doing so, we implicitly assume that the observations 
within an efficiency class are representative.} 

## Resampling the original data

To calibrate the sampling process information in the aggregate distribution of 
types (houses vs. apartments) and efficiency classes is required. Also the 
number of observations to be included in the final data set can be defined in 
this section.


```{r resampling, message=FALSE, warning=FALSE, results=FALSE}
# set size of final dataset
n_dataset <- 250000

# define the known distribution 
type_share <- c("Apartment" = 0.537, "House" = 0.463)
type_distribution <- type_share * n_dataset
efficiency_share <- data.frame(
          "efficiency_class" = c("+", "A", "B", "C", "D", "E", "F", "G", "H"),
          "realdist"=c(4, 3, 7, 12, 15, 14, 15, 13, 17)/100)

# create a representative sample
set.seed(1301)
immo_rep1 <- immo_raw %>%
  split(.$type) %>%
  imap_dfr(~slice_sample(.x, n = round(type_distribution[.y]), replace=TRUE)) %>%
  mutate(index=seq(1:n_dataset))

immo_rep1 <- immo_rep1[sample(nrow(immo_rep1)),]

immo_classes_rep1 <- immo_rep1 %>%
  group_by(efficiency_class) %>%
  summarise(buildings = sum(renovation_weight),
            observations = n(),  
            raw_class_share = n()/nrow(immo_rep1),
            weighted_class_share = sum(renovation_weight)/
                                        sum(immo_rep1$renovation_weight),
            mean_weight = mean(renovation_weight)) %>%
            mutate(realdist = efficiency_share$realdist,
            absdistance = weighted_class_share - realdist,
            reldistance = (weighted_class_share - realdist) / 
                                  realdist) %>%
  ungroup()
  
least_represented <- immo_classes_rep1 %>%
  filter(reldistance==min(immo_classes_rep1$reldistance))
  
immo_classes_rep1 <- immo_classes_rep1 %>%
                     mutate(todraw = floor(pull(least_represented, buildings)/
                            pull(least_represented, realdist)*realdist),
                            todraw_weighted = floor(todraw / mean_weight))

immo_rep1 <- immo_rep1 %>%
  group_by(efficiency_class) %>%
  mutate(cum_class_weight = cumsum(renovation_weight)) %>%
  ungroup() 

for (i in efficiency_share$efficiency_class){
sampled_data <- filter(immo_rep1, efficiency_class==i & cum_class_weight < pull(filter(immo_classes_rep1,efficiency_class==i),todraw))

 if(i=="+"){
    immo_rep <- sampled_data
  } else {
    immo_rep <- merge(immo_rep, sampled_data, all=TRUE)
  }
}

for (i in efficiency_share$efficiency_class){
  print(i)
  print(pull(filter(immo_classes_rep1,efficiency_class==i),todraw))
  print(max(pull(filter(immo_rep,efficiency_class==i),cum_class_weight)))
  print(sum(select(filter(immo_rep,efficiency_class==i),renovation_weight)))
}
  

```
## Plotting the representative dataset

```{r representativeplot, message=FALSE, warning=FALSE}
plot_rep <- ggplot(immo_rep, aes(x=ec_index)) +
  geom_density(aes(colour= efficiency_class), alpha=.2, fill="#FF6666", 
               show.legend = TRUE) +
  labs(title="All", x="ec_index", y="Density (accommodation units)") +
  xlim(0, 500) +
  ylim(0, 0.1) +
  theme_bw()

plot_rep

```

Export the plot as a pdf file.

```{r representativeplotPDF, message=FALSE, warning=FALSE}
ggsave("../Graphs/plot_rep.pdf", plot_rep, width = 6, height = 4)

```

## Plausibility checks
In this section we make us of some  [offical data](https://www.bmwk.de/Redaktion/DE/Downloads/Studien/vorbereitende-untersuchungen-zur-langfristigen-renovierungsstrategie-ergaenzung.pdf?__blob=publicationFile&v=6) to assess the properties of the representative sample. In detail we perform the 
following checks:

1. We first reproduce the share associated with building types and efficiency 
      classes used to calibrate the dataset.
2. On average, the thermal envelope of all private and commercial buildings in 
      Germany corresponds to energy efficiency class F, with an annual demand or 
      consumption of 170 kWh/m. With our data we can roughly reproduce this value.
3. We compare the share of squaremeters of houses and apartments in the worst 
      efficiency classes; these should be 40% (for houses) 
      and 16% (for apartments).

For the \textit{German case}, the first two test give convincing results, close 
to aggregate estimates. For the third test, we find that we cannot fully account 
for the fact that more badly insulated buildings are, on average, larger. 
Hence, our estimates.

```{r plausibility, message=FALSE, warning=FALSE}
# Confirm correct shares (1)
print(count(filter(immo_rep,type == "Apartment")) / nrow(immo_rep))
print(count(filter(immo_rep,type == "House")) / nrow(immo_rep))

immo_rep %>%
  group_by(efficiency_class) %>%
  summarise(share_effclasses = sum(renovation_weight)/
              sum(immo_rep$renovation_weight))     

# confirm average insulation quality (2)
weighted.mean(immo_rep$ec_index, immo_rep$renovation_weight)

# share of squaremeters in the worst efficiency-classes (3)
immo_rep %>%
  filter(type == "House") %>%
  group_by(efficiency_class) %>%
  summarise(share_effclasses = sum(living_area) / 
                    sum(immo_rep$living_area[immo_rep$type == "House"])) 

immo_rep %>%
  filter(type == "Apartment") %>%
  group_by(efficiency_class) %>%
  summarise(share_effclasses = sum(living_area) / 
                    sum(immo_rep$living_area[immo_rep$type == "Apartment"])) 
```
## Export the representative dataset

We export the dataset to save the cleaned data. This allows to remove temporary 
calculations in R, save calculation time, and import the cleaned data.

```{r exportrepdataset, message=FALSE, warning=FALSE, results=FALSE}
write.csv(immo_rep, here::here("Intermediate_Results/data_clean_representative.csv"), 
          row.names = FALSE)

```